---
title: "07-Cross-validation-missing-values"
author: "Andrew Balmer"
date: "2024-01-03"
output: html_document
---

```{r setup, include=FALSE}

# Clear the workspace
remove(list = ls())

# Load required packages
library(tidyverse)     # For data manipulation and visualization
library(smacof)        # For multidimensional scaling
library(RColorBrewer)  # For color palettes
library(calibrate)     # For calibration
library(matrixStats)   
library(ggExtra)       # For marginal histograms
library(ggdensity)     # For density distrbutions
library(foreach)
library(doParallel)

# Set the working directory
setwd("/Users/ajb306/AMR-cartography/analysis/#01-Phenotype_and_map_analyses/#13-19-Cross-validation-analyses/results/")

# Read the MIC table data
tablemic <- read.csv("/Users/ajb306/AMR-cartography/data/MIC_table_Spneumoniae.csv", header=TRUE, sep=",", skip = 0)

# Read in the relevant meta data
tablemic_meta <- read.csv("/Users/ajb306/AMR-cartography/data/meta_data_Spneumoniae.csv", header=TRUE, sep=",", skip = 0)

# Specify the full path to the data file
phen_map <- "/Users/ajb306/AMR-cartography/analysis/#01-Phenotype_and_map_analyses/#03-06-Generating-phenotype-and-genotype-maps/results/Spneumo_3628_PCA_start_2D_METRIC.RData"

# Load the pre-computed PCA start 2D metric data
load(phen_map)


```


### ==== Background

Measuring resistance phenotypes using MIC dilution series can be challenging to analyse. When the MIC assay is conducted for many isolates across different laboratories, it is not uncommon for individual MIC values to be missing in the final dataset. This can occur for a number of reasons, such limitations in cost, time or inability to grow an isolate on a given day. However, these missing values pose a challenge for downstream analysis, particularly when isolates have missing values for several drugs. During standard univariate analysis, these isolates are often simply excluded. However this is a considerable cost in the usefulness of the data and time spent collecting it. In contrast, with multivariate methods, the information present for the other drugs can be used to impute one or more missing MIC value.

One advantage of the MDS framework is its ability to take relatively imprecise data and generate a more accurate representation of that data. In theory, because isolates are positioned using several MIC correlated values, missing values for any individual can be predicted in the final representation. Here I tested the ability of MDS to predict missing MIC value. I did this by computationally removing MIC values (replaced with NA) and testing how well MDS could cope with this


## Aims and objectives 
In this document, I have used several methods to ensure MDS is robust to missing values in the underlying assay data. Firstly, I tried making maps with missing data for a subset of isolates and comparing them to the original maps made with their 'true' observed values. I then compared the two maps using the 'congruence coefficient' between the two maps. Secondly, I measured the euclidean distance between the true value on the map and its position with missing values. Lastly, I tested prediction error was increased or decreased when using more than two dimensions (a cross-validation analysis). 

I also calculated the levels of stress for the missing-value maps, as well as stress-per-point for the isolates with and without missing values. 

### Generating datasets with 'missing' values
In order to use these methods, I generated 100 datasets with missing value for 10% of the MIC entries (i.e. 10% of the 3628 * 6 MIC titres). A pairwise distance matrix for each dataset was then generated. In cases where one or both samples have missing values present, distances are calculated by excluding the column(s) with the missing value in estimating the pairwise distance to the other isolates, and estimating pairwise distance only using those present. In cases where MICs were included for all drugs for a pair of isolates (i,e. no missing values), all information was used to generate the distance estimate. 

# 1 - Generate Missing Sample Datasets
In this first chunk, I generate the 100 missing datasets needed to run the analysis. I do this by generating 100 copies of the original dataset, then randomly removing 10% of the MIC values from each. Lastly, I compute the distance matrices for both the original dataset, and the 100 'missing' sample datasets.

```{r, echo = F}

# Set seed 
set.seed(1234)

# Number of missing sample datasets to be run
no_missing_samples <- 8

# List to store missing sample datasets
missing_samples <- list()

# Generate missing samples by duplicating the original tablemic data
for (i in 1:no_missing_samples) {
  missing_samples[[i]] <- tablemic
  colnames(missing_samples[[i]]) <- colnames(tablemic)

  # Randomly delete a proportion (10%) of the data in each missing sample
  while (sum(is.na(missing_samples[[i]]) == TRUE) < (nrow(missing_samples[[i]]) * ncol(missing_samples[[i]]) * 10 / 100)) {
    missing_samples[[i]][sample(nrow(missing_samples[[i]]), 1), sample(ncol(missing_samples[[i]]), 1)] <- NA
  }
}

# Check the proportion of NA values in a missing sample to ensure the above worked
(sum(is.na(missing_samples[[1]])) / (nrow(tablemic) * ncol(tablemic))) * 100

# List to store distance matrices of missing samples
missing_samples_dists <- list()

# Compute distance matrices for each missing sample
for (i in 1:no_missing_samples) {
  missing_samples_dists[[i]] <- dist(missing_samples[[i]])
}

# Compute the distance matrix for the original data
dist_pne <- dist(tablemic)

```

# 2 - Reshape Data for Analysis

```{r,echo = F}
# List to store actual values of each missing sample
missing_samples_values <- list()

# Extract LABID, MIC values, and reshape the data for each missing sample
for (i in 1:no_missing_samples) {
  missing_samples_values[[i]] <- missing_samples[[i]]
  missing_samples[[i]] <- cbind(tablemic_meta[, 1], missing_samples[[i]])

  missing_samples[[i]] <- missing_samples[[i]] %>%
    rename(LABID = 1)

  missing_samples[[i]] <- missing_samples[[i]] %>% 
    pivot_longer(!LABID, names_to = "drug", values_to = "MIC_value")
}

```

# 3 - Merge Actual Values with Missing Samples

Here, we merge the missing value tables with the real values. This will be useful later, as this can be used to compare the true values to their predicted values

```{r, echo = F}
# Create a data frame with the actual values of the original data
tablemic_real_values <- cbind(tablemic_meta[, 1], tablemic)

tablemic_real_values <- tablemic_real_values %>% 
  rename(LABID = 1)

tablemic_real_values <- tablemic_real_values %>% pivot_longer(!LABID, names_to = "drug", values_to = "true_value")

# Merge actual values with missing samples and add a dataset column
for (i in 1:no_missing_samples) {
  missing_samples[[i]]$LABID <- as.character(missing_samples[[i]]$LABID)
  missing_samples[[i]] <- left_join(missing_samples[[i]], tablemic_real_values, by = c("LABID" = "LABID", "drug" = "drug")) %>%
    mutate(dataset = i)
}

```

# 4 - Compute Weight Matrices for Each Missing Sample

Isolates with one or more missing value were also given less weight in generating the map. This lessens the effect of isolates with missing MIC values on the positioning of other isolates. This makes sense as isolates without missing values have more information on where to position them. To do this, a weight matrix was constructed for each of the 100 samples based on how may MIC values were missing for each isolate. To generate this weight matrix, the pairwise distance between each pair of isolates was weighted by multiplying the number of MIC values present for that pair of isolates, then squaring the result  The absolute values of the weight matrix is less important than the differences between them. Under this weighting structure, an isolate with 3 missing values should therefore be weighted less strongly than one with 2, and 2 would be less strongly than 1 and so on. Finding the precise weighting structure with the least prediction error is a process of trial and error, but this structure seems to work well for the S. pneumoniae map. The MDS algorithm is then used later to generate a map for each of the 100 samples, using the metric, PCoA start method.

```{r, echo = F}

# List to store weight matrices for each missing sample
weight_missing_samples <- list()

# List to store weight matrices for each missing sample
missing_samples_weight_matrices <- list()

# List to store column names of each missing sample
number_of_MIC_values <- list()

# Compute weight matrices for each missing sample based on actual values
for (i in 1:no_missing_samples) {
  number_of_MIC_values <- rowCounts(as.matrix(missing_samples_values[[i]]), value = NA, na.rm = FALSE)
  number_of_MIC_values <- (number_of_MIC_values * -1)
  number_of_MIC_values <- number_of_MIC_values + -min(number_of_MIC_values, na.rm = TRUE) + 1

  missing_samples_weight_matrices[[i]] <- matrix(1, nrow(tablemic), nrow(tablemic))

  for (f in 1:ncol(dist_pne)) {
    missing_samples_weight_matrices[[i]][, f] <- number_of_MIC_values * missing_samples_weight_matrices[[i]][, f] 
  }

  missing_samples_weight_matrices[[i]] <- t(missing_samples_weight_matrices[[i]])

  for (f in 1:ncol(dist_pne)) {
    missing_samples_weight_matrices[[i]][, f] <- number_of_MIC_values * missing_samples_weight_matrices[[i]][, f] 
  }

  diag(missing_samples_weight_matrices[[i]]) = 0
}

# Square each element of the weight matrices so as to weight samples with more MIC values more strongly
for (i in 1:no_missing_samples) {
  missing_samples_weight_matrices[[i]] <- missing_samples_weight_matrices[[i]] * missing_samples_weight_matrices[[i]]
}

# Extract LABID and MIC values for each missing sample
for (i in 1:no_missing_samples) {
  missing_samples_values[[i]] <- cbind(tablemic_meta[, 1], missing_samples_values[[i]])
  missing_samples_values[[i]] <- missing_samples_values[[i]] %>% 
    rename(LABID = 1)
}

```


# 5 - Run MDS on each of the 100 datasets with missing values

Next we run MDS on each of the 100 datasets with missing values. In this case, the 100 datasets are split into four batches and processed in parallel to speed up computation. 

```{r, echo = F}

missing_samples_mds_objects <- NULL

# Set the number of cores to use
num_cores <- 4  # Change this to the desired number of cores

# Register parallel backend
registerDoParallel(cores = num_cores)

# Define ranges for each core
core_ranges <- rep(ceiling(no_missing_samples / num_cores), num_cores)

# Define a function to perform MDS on a chunk of datasets
run_mds <- function(indices) {
  missing_values_test <- list()
  
  for (i in 1:no_missing_samples) {
    temp_MDS <- mds(missing_samples_dists[[i]], ndim = 2, type = c("ratio"), 
                    weightmat = missing_samples_weight_matrices[[i]], init = "torgerson", 
                    modulus = 1, itmax = 1000, eps = 1e-06)
    
    missing_values_test[[i]] <- list(conf = temp_MDS$conf, stress = temp_MDS$stress, spp = temp_MDS$spp)
    temp_MDS <- NULL
  }
  
  return(missing_values_test)
}

# Run MDS in parallel
missing_samples_mds_objects <- foreach(indices = core_ranges, .combine = "c") %dopar% {
  run_mds(1:no_missing_samples)
}

# Stop parallel processing
registerDoSEQ()


# Save and restore the MDS results
saveRDS(missing_samples_mds_objects, file = "missing_samples_mds_objects.rds")
missing_samples_mds_objects <- readRDS(file = "missing_samples_mds_objects.rds")


```

# 6 - Saving the generated objects
As the objects generated by the previous processes are quite large, I have also included code to save each of these. 

```{r, echo = F}

# Define a list of objects to save
objects_to_save <- list(
  missing_samples = missing_samples,
  missing_samples_dists = missing_samples_dists,
  missing_samples_weight_matrices = missing_samples_weight_matrices,
  missing_samples_values = missing_samples_values#,
  #missing_samples_mds_objects = missing_samples_mds_objects
)

# Loop to save each object
for (obj_name in names(objects_to_save)) {
  saveRDS(objects_to_save[[obj_name]], file = paste0(obj_name, ".rds"))
}

# Loop to restore each object
for (obj_name in names(objects_to_save)) {
  assign(obj_name, readRDS(file = paste0(obj_name, ".rds")))
}

```

# Load the S. pneumoniae phenotype map

Here, I load in the S. pneumoniae phenotype map object and rotate/dilate it to an interpretable scale. 

```{r, echo =F}
# Set the rotation angle in radians
theta <- 326 * pi / 180

# Create a rotation matrix
rot <- matrix(c(cos(theta), sin(theta), -sin(theta), cos(theta)), ncol = 2)

# Apply rotation to the configuration data
torg_met$conf <- torg_met$conf %*% rot

# Initialize a list to store rotated configurations of missing samples
missing_samples_dists_confs <- list()

# Extract rotated configurations from each missing sample MDS result
for (i in 1:no_missing_samples) {
  missing_samples_dists_confs[[i]] <- as_tibble(missing_samples_mds_objects[[i]]$conf)
}

# Initialize a vector to store stress values of missing samples
missing_samples_stress <- vector()

# Extract stress values from each missing sample MDS result
for (i in 1:no_missing_samples) {
  missing_samples_stress[i] <- missing_samples_mds_objects[[i]]$stress
}

# Convert the stress vector to a tibble
missing_samples_stress <- as_tibble(missing_samples_stress)

# Extract original configurations from the "torg_met" object
torg_met_conf <- as.data.frame(torg_met$conf)

# Extract distance matrices from "torg_met"
table_distances <- as_tibble(as.matrix(torg_met[[1]]))
map_distances <- as_tibble(as.matrix(torg_met[[3]]))

# Rename columns of map_distances to match table_distances
colnames(map_distances) <- colnames(table_distances)

# Reshape distance matrices into long format using pivot_longer
table_distances <- pivot_longer(table_distances, cols = 1:nrow(tablemic), names_to = "antibiotic", values_to = "table_distance")
map_distances <- pivot_longer(map_distances, cols = 1:nrow(tablemic), names_to = "antibiotic", values_to = "map_distance")

# Combine distance values into a single data frame
distances <- bind_cols(table_distances, map_distances)

# Convert distance values to numeric
distances$table_distance <- as.numeric(distances$table_distance)
distances$map_distance <- as.numeric(distances$map_distance)

# Fit a linear model to relate map_distance to table_distance
mapvtable <- lm(map_distance ~ table_distance, data = distances)

# Extract slope from the linear model
slope <- as.numeric(coef(mapvtable)[2])

# Calculate the dilation factor based on the slope
dilation <- 1 / slope

# Rename columns of the original configuration data
colnames(torg_met_conf) <- c("D1", "D2")

# Apply dilation to each column of the original configuration data
for (i in 1:ncol(torg_met_conf)) {
  torg_met_conf[, i] <- torg_met_conf[, i] * dilation
}


```


### Congruence coefficient.

To more quantitatively compare the full map with those with missing values, I used procrustes to rotate and dilate the maps into the same rotation. I first calculated the congruence coefficient between the sets of maps which can be used too compare the similarity of two solutions (https://en.wikipedia.org/wiki/Congruence_coefficient#:~:text=In%20multivariate%20statistics%2C%20the%20congruence,derived%20in%20a%20factor%20analysis.&text=It%20can%20be%20used%20to,have%20taken%20the%20same%20test.)

Below, the plot shows that the maps with missing values are essentially the same as the one made without these isolates. This suggests the missing values are not having a strong effect on how the maps are positioning the isolates with actual values. Generally, a congruence coefficient above 0.95 is considered very similar, and above .99 is near identical For each transformation. hHre, the estimate is >0.998, suggesting the solutions are essentially the same. While there are more fine-grain measures to measure error for the isolates with missing values (see below), this parameter provides a good estimate of the overall similarity of the maps.

```{r, echo = F}

# Initialize variables to store intermediate results
mapvtable <- NULL
map_distances <- NULL
table_distances <- NULL
distances <- NULL

# Initialize lists to store Procrustes analysis results
met_ord_comparison <- list()
met_ord_comparison_congcoef <- vector()
met_ord_comparison_aliencoef <- vector()

# Perform Procrustes analysis for each missing sample
for (i in 1:no_missing_samples) {
  met_ord_comparison[[i]] <- Procrustes(as.matrix(torg_met_conf), as.matrix(missing_samples_dists_confs[[i]]))
  met_ord_comparison_congcoef[i] <- met_ord_comparison[[i]]$congcoef
  met_ord_comparison_aliencoef[i] <- met_ord_comparison[[i]]$aliencoef
}

# Convert alignment coefficients to a tibble
met_ord_comparison_congcoef <- as_tibble(met_ord_comparison_congcoef)

# Calculate and print the mean of congruence coefficients
mean_congcoef <- round(mean(met_ord_comparison_congcoef$value), 3)
cat("Mean Congruence Coefficient:", mean_congcoef, "\n")

# Plot histogram of congruence coefficients
B <- ggplot(met_ord_comparison_congcoef, aes(x = value)) + 
  geom_histogram(position = "identity", alpha = 0.5, color = "black", fill = "#E41A1C", bins = 30) + 
  geom_vline(aes(xintercept = mean(value)),
             color = "#E41A1C", linetype = "dashed", size = 1) + 
  geom_vline(aes(xintercept = 0.99),
             color = "black", linetype = "solid", size = 1) +
  theme_bw() +
  labs(x = "Congruence Coefficient", y = "Count") +
  coord_fixed(ratio = 0.000075) 
B
# Save the plot as a JPEG file
ggsave("S_pneumoniae_missing_cong_conf.jpg", B)


```


### Testing whether datasets with additional error have higher stress

In MDS, stress is the sum of the squared distances between the points on the plot and their distances in the original data table. This is the parameter MDS tries to minimise when making the map. Theoretically, given there strong correlations between MICs to the different drugs, adding missing values might effect their stress values on the map. This is because the missing values may have the effect of altering those correlations, and meaning the distances can not be as well represented in 2 dimensions. 

I estimated stress for each of the maps with missing values (histogram) and compared them to the stress for the original map (black line). Notably, all of the missing value maps had higher stress than the original map. Moreover, the isolates which had missing values tended to have a higher level of stress-per-point (as a % of total stress). Below is the histogram of stress per point for the missing values (red) compared to the other isolates on the map (grey). The stress for the points with missing values is slightly larger than those without.

Notably, some isolates had missing values for more than one of their MIC values (up to a maximum of 5 of the 6 MIC values), below I have aggregated this error in this plot. Isolates contribution to stress was higher when they had multiple missing values. 


```{r, echo = F}

# Initialize variables to store intermediate results
missing_samples_stress <- vector()
missing_samples_stress_torg <- vector()

# Extract pairwise distances and additional information for each missing sample
for (i in 1:no_missing_samples) {
  met_ord_comparison[[i]]$pairwise_dist <- cbind(met_ord_comparison[[i]]$X, met_ord_comparison[[i]]$Yhat)
  colnames(met_ord_comparison[[i]]$pairwise_dist) <- c("D1_X", "D2_X", "D1_Y", "D2_Y")

  met_ord_comparison[[i]]$pairwise_dist <- as_tibble(met_ord_comparison[[i]]$pairwise_dist) %>% 
    mutate(dist_phen = sqrt((D1_X - D1_Y)^2 + (D2_X - D2_Y)^2),
           LABID = tablemic_meta[, 1])

  missing_samples_mds_objects[[i]]$spp <- cbind(tablemic_meta$LABID, missing_samples_mds_objects[[i]]$spp)
  colnames(missing_samples_mds_objects[[i]]$spp) <- c("LABID", "spp")

  met_ord_comparison[[i]]$pairwise_dist <- left_join(met_ord_comparison[[i]]$pairwise_dist, 
                                                     as_tibble(missing_samples_mds_objects[[i]]$spp), 
                                                     by = "LABID")
}

# Combine pairwise distances and additional information
torg_met_conf <- cbind(torg_met_conf, tablemic_meta$LABID, torg_met$spp)
colnames(torg_met_conf) <- c("D1", "D2", "LABID", "stress_per_point_real")

# Initialize a list to store stress per point comparison results
stress_per_point_comparison <- list()

# Perform additional data manipulations and comparisons
for (i in 1:no_missing_samples) {
  stress_per_point_comparison[[i]] <- left_join(met_ord_comparison[[i]]$pairwise_dist, missing_samples[[i]], by = "LABID")
  stress_per_point_comparison[[i]]$spp <- as.numeric(stress_per_point_comparison[[i]]$spp)

  stress_per_point_comparison_2 <- left_join(stress_per_point_comparison[[i]], torg_met_conf, by = "LABID") %>%
    group_by(LABID) %>%
    mutate(count_na = sum(is.na(MIC_value))) %>%
    ungroup() %>%  
    distinct(LABID, .keep_all = TRUE) %>%
    dplyr::select(LABID, count_na, stress_per_point_real) 

  stress_per_point_comparison[[i]] <- left_join(stress_per_point_comparison[[i]], stress_per_point_comparison_2, by = c("LABID" = "LABID")) 

  stress_per_point_comparison[[i]]$spp <- as.numeric(stress_per_point_comparison[[i]]$spp)

  stress_per_point_comparison[[i]] <- stress_per_point_comparison[[i]] %>%
    mutate(dataset = i)
}

# Combine results into a single data frame and add a column 'missing_values' based on the count of missing values
stress_per_point_comparison <- bind_rows(stress_per_point_comparison) %>% 
  mutate(missing_values = ifelse(count_na == "0", "No missing values", "Missing values"))


# Calculate and print the difference in means of spp values
mu <- stress_per_point_comparison %>%
  group_by(missing_values) %>%
  summarize(Mean = mean(spp, na.rm = TRUE))

difference_in_means <- round(mu$Mean[1] - mu$Mean[2], 3)
cat("Difference in Means:", difference_in_means, "\n")

# Plot histogram of overall map stress
ggplot(as_tibble(missing_samples_stress), aes(x = value)) + 
  geom_histogram(color = "black", alpha = 0.5, fill = "#E41A1C", bins = 20) +
  theme_bw() +
  geom_vline(xintercept = torg_met$stress) +
  labs(x = "Overall map stress", y = "Count") 

# Plot histogram and density plot of stress per point
E <- ggplot(stress_per_point_comparison, aes(spp, y = stat(density), colour = missing_values, fill = missing_values)) + 
  geom_histogram(position = "identity", alpha = 0.45, color = "black") + 
  geom_freqpoly(bins = 30, stat = "bin") + 
  theme_bw() + 
  geom_vline(data = mu, aes(xintercept = Mean, color = missing_values),
             linetype = "dashed", size = 0.75) +
  scale_x_continuous(breaks = seq(0, max(abs(stress_per_point_comparison$spp), na.rm = TRUE) + 0.2, 0.2)) +
  scale_y_continuous(breaks = seq(0, 60, 10)) +
  labs(x = "Stress per point (%)", y = "Density") +
  theme(legend.position = "none") + 
  guides(colour = "none") +
  guides(fill = guide_legend(title = "Error added")) +
  scale_colour_manual(values = c("#E41A1C", "darkgrey")) +
  scale_fill_manual(values = c("#E41A1C", "darkgrey"))
E

# Save the plot as a JPEG file
ggsave("S_pneumo_missing_spp.jpg", E)


```


### Euclidean distance between true values and predicted values

After reorientation of the missing value maps to compare to the real map, I then estimated the euclidean distance between each isolates true value and their missing value. Below is a histogram of those values for the isolates with missing MICs (Red) and the distances for those without noise added (grey). Typically the values were less than 1 MIC unit away from their true values. Generally the distances tended to be slightly higher when multiple MICs were missing. However, some points were very far away from their true values. This was typically those which had error added for several drugs or high very high contribution to stress on the map with or without the missing value.


```{r, echo = F}

# Calculate and print the mean of dist_phen values
mu <- stress_per_point_comparison %>%
  group_by(missing_values) %>%
  summarize(Mean = mean(dist_phen, na.rm = TRUE))

mean_dist_phen <- round(mu$Mean[1], 3)
cat("Mean Distance from True Value:", mean_dist_phen, "\n")

# Plot histogram and density plot of dist_phen
C <- ggplot(filter(stress_per_point_comparison), aes(dist_phen, y = stat(density), colour = missing_values, fill = missing_values)) + 
  geom_histogram(position = "identity", alpha = 0.45, color = "black") + 
  geom_freqpoly(bins = 40, stat = "bin") + 
  theme_bw() + 
  geom_vline(data = mu, aes(xintercept = Mean, color = missing_values),
            linetype = "dashed", size = 0.75) +
  geom_vline(aes(xintercept = 1), color = "black",
             linetype = "solid", size = 0.5) +
  scale_x_continuous(breaks = seq(0, 7, 1)) +
  labs(x = "Distance from true value (MIC units)", y = "Density") +
  theme(legend.justification = c(0.95, 0.725), legend.position = c(0.95, 0.8),
        axis.text = element_text(size = 16), 
        axis.title = element_text(size = 16)) +
  guides(colour = "none") +
  guides(fill = guide_legend(title = "Error added")) +
  scale_colour_manual(values = c("#E41A1C", "darkgrey")) +
  scale_fill_manual(values = c("#E41A1C", "darkgrey")) +
  theme(legend.position = "none") 
C
# Save the plot as a JPEG file
ggsave("S_pneumoniae_noise_euc_phen_dist.jpg", C)


```



### Visually comparing a noise-added map to true values

I then plotted a map made with missing values and the original map on the same plot after reorienting them to the same scale/rotation. Here, the black points are the true values, while the red points are the noise added values. The black crosses indicate points which did not have error added. Although it is a little difficult to see because of the number of points, most values are within 1log fold of their true value (see above).

```{r, echo = F}


# Create a scatter plot with segments for missing values
A <- ggplot(filter(stress_per_point_comparison, missing_values != "No missing values" & dataset == 1), aes(x = D1_X, y = D2_X)) + 
  geom_point(data = filter(stress_per_point_comparison, missing_values == "No missing values" & dataset == 1), aes(x = D1_X, y = D2_X), shape = 4, fill = "grey", size = 1.5, alpha = 0.6) +
  geom_segment(data = filter(stress_per_point_comparison, missing_values != "No missing values" & dataset == 1), aes(x = D1_X, y = D2_X, xend = D1_Y, yend = D2_Y), size = 0.5, colour = "grey") + 
  geom_point(data = filter(stress_per_point_comparison, missing_values != "No missing values" & dataset == 1), shape = 21, fill = "black", size = 2, colour = "white") + 
  geom_point(data = filter(stress_per_point_comparison, missing_values != "No missing values" & dataset == 1), aes(x = D1_Y, y = D2_Y), shape = 21, fill = "#E41A1C", size = 2, colour = "white") +
  theme_bw() +   
  scale_x_continuous(breaks = seq(min(c(stress_per_point_comparison$D1_Y, stress_per_point_comparison$D1_X), na.rm = TRUE), max(c(stress_per_point_comparison$D1_Y, stress_per_point_comparison$D1_X), na.rm = TRUE) + 1, 1)) +
  scale_y_continuous(breaks = seq(min(c(stress_per_point_comparison$D2_Y, stress_per_point_comparison$D2_X), na.rm = TRUE), max(c(stress_per_point_comparison$D2_Y, stress_per_point_comparison$D2_X), na.rm = TRUE) + 1, 1)) +
  labs(title = "") +
  labs(x = "MDR distance", y = "MDR distance") + 	
  theme(axis.text = element_text(size = 16), 
        axis.title = element_text(size = 16), 
        panel.grid.major = element_line(colour = "grey", size = 0.3),
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.text.x = element_text(size = 14)) +
  coord_fixed()
A
# Save the plot as a JPEG file
ggsave("S_pneumo_noise_pred_error_map.jpg", A)

```




### Cross-validation analysis - multiple dimensions

Lastly, I made maps in several different dimensions using the same datasets with missing values as before. I also made maps using the true values for each dimension to compare them to. Here, I again calculated the euclidean distance between the 'true' map position and the missing value map position, after reorientation of the maps to overlap. I found that prediction error was lowest at 1 dimension, but that there were marginal increases in error when using more than 2 dimensions. This suggests that using more than two dimensions may overfit the data. Although 1 dimension offered the lowest prediction error here, it was clear from the stress and goodness of fit analyses (previous markdowns) that a minimum of two dimensions was needed to accurately represent the data. 

Here, I plotted each dimension on the x axis, and the mean prediction error for each missing value dataset on the y axis. The points and error bars show the standard error for the mean of the means for each dataset.


```{r, echo = FALSE}

# Initialize a list to store results for each dimension
missing_samples_dists_dimensions <- list()

# Create nested lists for each dimension
for (f in 1:4){
  missing_samples_dists_dimensions[[f]] <- list()
}

# Set the number of samples
no_missing_samples <- 1

# Set the number of cores to use
num_cores <- 4  # Change this to the desired number of cores

# Register parallel backend
registerDoParallel(cores = num_cores)

# Function to perform MDS for a given dimension
run_mds_dimension <- function(dim) {
  MDS_missing_samples_dim_test <- list()

  # Loop over samples
  for (i in 1:no_missing_samples) {
    # Perform MDS
    temp_MDS <- mds(missing_samples_dists[[i]], ndim = dim, type = c("ratio"),
                    weightmat = missing_samples_weight_matrices[[i]], init = "torgerson",
                    modulus = 1, itmax = 1000, eps = 1e-06)

    # Store MDS results
    MDS_missing_samples_dim_test[[dim]][[i]] <- list(conf = temp_MDS$conf, stress = temp_MDS$stress, spp = temp_MDS$spp)
    temp_MDS <- NULL
  }
  
  return(MDS_missing_samples_dim_test)
}
26/6
# Run MDS for each dimension in parallel
missing_samples_dists_dimensions <- foreach(dim = 1:4, .combine = "c") %dopar% {
  run_mds_dimension(dim)
}

# Stop parallel processing
registerDoSEQ()

# Save and restore the MDS results
saveRDS(missing_samples_dists_dimensions, file = "missing_samples_dists_dimensions_missing_values.rds")
missing_samples_dists_dimensions <- readRDS(file = "missing_samples_dists_dimensions_missing_values.rds")



```


```{r, echo = F}
# generate maps in several dimensions for all data in order to compare
torg_met_dimensions <- list()

for (f in 1:4){
torg_met_dimensions[[f]] <- mds(dist_pne, ndim = f, type = c("ratio") , init = "torgerson", modulus = 1, itmax = 1000, eps = 1e-06)
}

# Save an object to a file
saveRDS(torg_met_dimensions, file = "torg_met_dimensions_missing_values.rds")
# Restore the object
torg_met_dimensions <- readRDS(file = "torg_met_dimensions_missing_values.rds")


```


```{r, echo = F}

# Initialize lists for storing results
missing_samples_dists_confs_dim <- list()
missing_samples_stress_dim <- list()
torg_met_dimensions_conf <- list()
torg_met_dimensions_stress <- vector()

# Loop over dimensions
for (f in 1:4){
  missing_samples_dists_confs_dim[[f]] <- list()
  missing_samples_stress_dim[[f]] <- vector()

  # Loop over samples
  for (i in 1:no_missing_samples) {
    missing_samples_dists_confs_dim[[f]][[i]] <- as_tibble(missing_samples_dists_dimensions[[f]][[i]]$conf)
    missing_samples_stress_dim[[f]][i] <- missing_samples_dists_dimensions[[f]][[i]]$stress
  }

  missing_samples_stress_dim[[f]] <- as_tibble(missing_samples_stress_dim[[f]])

  torg_met_dimensions_conf[[f]] <- as_tibble(torg_met_dimensions[[f]]$conf)
  torg_met_dimensions_stress[f] <- torg_met_dimensions[[f]]$stress
}

torg_met_dimensions_stress <- as_tibble(torg_met_dimensions_stress)

# Initialize vectors for storing results
slope_dim <- vector()
dilation_dim <- vector()

# Loop over dimensions
for (f in 1:4) {
  # Process distances
  table_distances <- as_tibble(as.matrix(torg_met_dimensions[[f]][[1]]))
  map_distances <- as_tibble(as.matrix(torg_met_dimensions[[f]][[3]]))
  colnames(map_distances) <- colnames(table_distances)
  table_distances <- gather(table_distances, "antibiotic", "table_distance", 1:nrow(tablemic)) 
  map_distances <- gather(map_distances, "antibiotic", "map_distance", 1:nrow(tablemic))
  distances <- bind_cols(table_distances, map_distances)
  distances$table_distance <- as.numeric(distances$table_distance)
  distances$map_distance <- as.numeric(distances$map_distance)
  
  # Fit linear model
  mapvtable <- lm(map_distance ~ table_distance, data = distances)
  summary(mapvtable)
  mapvtable
  coef(mapvtable)
  
  # Store results
  slope_dim[f] <- as.numeric(coef(mapvtable)[2])
  dilation_dim[f] <- 1/slope_dim[f]
}

# Apply dilation to each dimension separately
for (f in 1:4){
for (i in 1:ncol(torg_met_dimensions_conf[[f]])){
  torg_met_dimensions_conf[[f]][,i] <- torg_met_dimensions_conf[[f]][,i] * dilation_dim[f]
}
}

# Initialize lists for storing results
met_ord_comparison_dim <- list()
met_ord_comparison_congcoef_dim <- list()

# Loop over dimensions
for (f in 1:4){
  met_ord_comparison_dim[[f]] <- list()
}

# Loop over dimensions and samples
for (f in 1:4) {
  for (i in 1:no_missing_samples) {
    # Perform Procrustes analysis
    met_ord_comparison_dim[[f]][[i]] <- Procrustes(as.matrix(torg_met_dimensions_conf[[f]]),  as.matrix(missing_samples_dists_confs_dim[[f]][[i]]))
  }
}

# Loop over dimensions and samples
for (f in 1:4) {
  for (i in 1:no_missing_samples) { 
    colnames(met_ord_comparison_dim[[f]][[i]]$X) <- paste(colnames(met_ord_comparison_dim[[f]][[i]]$X),"X",sep="_")
    colnames(met_ord_comparison_dim[[f]][[i]]$Yhat) <- paste(colnames(met_ord_comparison_dim[[f]][[i]]$Yhat),"Y",sep="_")
  }
}

# Loop over dimensions and samples
for (f in 1:4) {
  for (i in 1:no_missing_samples) {
    met_ord_comparison_dim[[f]][[i]]$pairwise_dist <- cbind(met_ord_comparison_dim[[f]][[i]]$X, met_ord_comparison_dim[[f]][[i]]$Yhat)
  }
}

# Loop over dimensions and samples
for (f in 1:4) {
  for (i in 1:no_missing_samples) {
    met_ord_comparison_dim[[f]][[i]]$pairwise_dist <- as_tibble(met_ord_comparison_dim[[f]][[i]]$pairwise_dist)
  }
}

# Loop over samples
for (i in 1:no_missing_samples) {
  met_ord_comparison_dim[[1]][[i]]$pairwise_dist <- as_tibble(met_ord_comparison_dim[[1]][[i]]$pairwise_dist) %>%
       mutate(dist_phen = sqrt((D1_X-D1_Y)^2),
              LABID = tablemic_meta[,1])
}

# Loop over samples
for (i in 1:no_missing_samples) {
  met_ord_comparison_dim[[2]][[i]]$pairwise_dist <- as_tibble(met_ord_comparison_dim[[2]][[i]]$pairwise_dist) %>%
       mutate(dist_phen = sqrt((D1_X-D1_Y)^2 + (D2_X-D2_Y)^2),
              LABID = tablemic_meta[,1])
}

# Loop over samples
for (i in 1:no_missing_samples) {
  met_ord_comparison_dim[[3]][[i]]$pairwise_dist <- as_tibble(met_ord_comparison_dim[[3]][[i]]$pairwise_dist) %>%
       mutate(dist_phen = sqrt((D1_X-D1_Y)^2 + (D2_X-D2_Y)^2 + (D3_X-D3_Y)^2),
              LABID = tablemic_meta[,1])
}

# Loop over samples
for (i in 1:no_missing_samples) {
  met_ord_comparison_dim[[4]][[i]]$pairwise_dist <- as_tibble(met_ord_comparison_dim[[4]][[i]]$pairwise_dist) %>%
       mutate(dist_phen = sqrt((D1_X-D1_Y)^2 + (D2_X-D2_Y)^2 + (D3_X-D3_Y)^2 + (D4_X-D4_Y)^2),
              LABID = tablemic_meta[,1])
}

# Loop over dimensions
for (f in 1:4) {
  # Loop over samples
  for (i in 1:no_missing_samples) {
    missing_samples_dists_dimensions[[f]][[i]]$spp <- cbind(tablemic_meta$LABID, missing_samples_dists_dimensions[[f]][[i]]$spp)
    colnames(missing_samples_dists_dimensions[[f]][[i]]$spp) <- c("LABID", "spp")
    
    met_ord_comparison_dim[[f]][[i]]$pairwise_dist <- left_join(met_ord_comparison_dim[[f]][[i]]$pairwise_dist, as_tibble(missing_samples_dists_dimensions[[f]][[i]]$spp), by = "LABID")
  }
}

# Initialize lists for storing results
stress_per_point_comparison <- list()
stress_per_point_comparison_2 <- list()

# Loop over dimensions
for (f in 1:4){
  stress_per_point_comparison[[f]] <- list()
  stress_per_point_comparison_2[[f]] <- list()

# Loop over dimensions and samples
  for (i in 1:no_missing_samples) {
    stress_per_point_comparison[[f]][[i]] <- left_join(as_tibble(met_ord_comparison_dim[[f]][[i]]$pairwise_dist), as_tibble(missing_samples[[i]]), by = "LABID")

# Loop to calculate count of missing values for each LABID in each dimension and dataset
    stress_per_point_comparison_2[[f]][[i]] <- stress_per_point_comparison[[f]][[i]] %>%
      group_by(LABID) %>%
      summarize(count_na = sum(is.na(MIC_value)))

    stress_per_point_comparison[[f]][[i]] <- left_join(stress_per_point_comparison[[f]][[i]], stress_per_point_comparison_2[[f]][[i]], by = c("LABID" = "LABID")) 
    
    stress_per_point_comparison[[f]][[i]] <- stress_per_point_comparison[[f]][[i]] %>%
      mutate(dimension = f, 
             dataset = i)
  }
}

# Combine all lists into a single dataframe
stress_per_point_comparison <- bind_rows(stress_per_point_comparison)

# Add a column indicating if there are missing values or not
stress_per_point_comparison <- stress_per_point_comparison %>% mutate(missing_values = ifelse(count_na == "0", "No missing values", "Missing values")) 

# Convert dist_phen column to numeric
stress_per_point_comparison$dist_phen <- as.numeric(stress_per_point_comparison$dist_phen)

# Convert dimension column to factor
stress_per_point_comparison$dimension <- as.factor(stress_per_point_comparison$dimension)

# Calculate mean and median for observations with missing values
mu <- stress_per_point_comparison %>%
  filter(missing_values == "Missing values") %>%
  distinct(LABID, dimension, dataset, .keep_all = TRUE) %>%
  ungroup() %>%
  group_by(dimension, dataset) %>%
  summarize(Mean = mean(dist_phen),
            Median = median(dist_phen)) %>%
  group_by(dimension) %>%
  summarize(sd = round(sd(Mean), 3),
            Mean = round(mean(Mean), 3),
            Median = round(median(Median), 3))

# Display summary statistics
t(mu) 

# Create a ggplot for visualization
ggplot(mu, aes(x = dimension, y = Mean)) + 
  geom_point(color = "#E41A1C") +
  geom_line(color = "#E41A1C") +
  geom_errorbar(aes(ymin = Mean - sd, ymax = Mean + sd), width = 0.2,
                position = position_dodge(0.05), color = "#E41A1C") +
  theme_bw() +
  theme(axis.text = element_text(size = 18), 
        axis.title = element_text(size = 18)) +
  labs(x = "Dimension", y = "Mean Prediction Error (MIC units)") +
  coord_fixed(ratio = 20) 

# Save the ggplot as an image
ggsave("S_pneumo_missing_values_cross_validation.jpg")

```
